---
title: "Stats 184 Final Project"
author: "Zhejing Shi, Jianzhuo, Liu"
output: html_notebook
---

<br>
<br>
<br>

#### **Guiding Question 1 : Does higher freedom score means higher happiness score?**
#### **Guiding Question 2: Do countries with higher Economy score tend to have higher happiness score?**
#### **Guiding Question 3: Do Life.Expectancy and happiness score have a positive relationship?**
<br>

#### We choose this topic because we are very interested in which countries have the high happiness scores, and why those countries have high scores.
<br>

#### One of the challenge is that we did not know linear regression before.However, we think it is important to apply such a method in our project. Thus, we learned it and applied it in our report.

<br> 

#### **Important findings of our investigation**
#### 1) North America region has the highest average happiness score
#### 2) Countries with higher freedom score seem to have higher happiness score
#### 3) Countries with higher Economy score tend to have higher happiness score
#### 4) Life.Expectancy score and happiness score have a positive relationship?

<br> 

#### Become acquainted with the data sources:
##### 1) For the first dataset "2017.csv", it is a World Happiness Report survey of collected in 2017. This dataset is originally released at the United Nations at an event celebrating International Day of Happiness on March 20th. 
##### 2) There are 155 different cases, and ach case represent a country in the world.
##### 3) The second dataset "counntry" is the country statistics originally comes from UN statistical databases. Each case represent a country 
##### 5) We plan to use 25 variables

<br>
<br>

### Exploration!!!
#### 1) Loadind the first dataset "Happy" that contains the happiness score for each country
```{r}
# all packages used for the assignment
getwd()
setwd('~/Desktop/Stats462/')
library(mosaicData)
library(ggplot2)
library(mosaic)

Happy = read.csv('2017.csv', header = T)
head(Happy)

### change the varible name so we can easily inner join with another dataset
names(Happy)[1] = "country"
names(Happy)[6] = 'Economy'
names(Happy)[8] = 'Life.Expectancy'
names(Happy)[11] = 'Government.Corruption'
head(Happy)

```
<br>
<br>
<br>
<br>

### 2) Loading the second dataset and join the two datasets.
##### *(Country Statistics originally comes from UN statistical databases. Each case represent a country)* 
```{r}

### The second Dataset containing countries information around the world
Country <- read.csv('country_profile_variables.csv', header = T)
head(Country)

### Change the variable names
names(Country)[3] = 'Surface.area'
names(Country)[4] = "Population"
names(Country)[5] = "Population.density"
names(Country)[6] = "Sex.ratio"
names(Country)[7] = "GDP.Gross.domestic.product"
names(Country)[8] = "GDP.growth.rate"
names(Country)[names(Country)=="Economy..Services.and.other.activity....of.GVA."] =  "Services.and.other.activity"
names(Country)[names(Country)=="Mobile.cellular.subscriptions..per.100.inhabitants."] =  "cellular.subscriptions"
names(Country)[names(Country)=="Education..Government.expenditure....of.GDP."] =  "Education.expenditure"
names(Country)[names(Country)=="Individuals.using.the.Internet..per.100.inhabitants."] =  "using.the.Internet"
names(Country)[names(Country)=="Fertility.rate..total..live.births.per.woman."] =  "Fertility.rate"
names(Country)[names(Country)=="Unemployment....of.labour.force."] =  "Unemployment"


### Select some of the variables to form a new country dataset
Country_New = Country[c("country", "Region", 'Surface.area', "Population", "Population.density", "Sex.ratio", "GDP.Gross.domestic.product",  "GDP.growth.rate", "Services.and.other.activity", "cellular.subscriptions","Education.expenditure", "using.the.Internet","Fertility.rate", "Unemployment")]

head(Country_New)


### Check distinct regions in the dataset
categories <- unique(Country_New$Region) 
categories


### Join the two datasets we have
Df<- merge(Happy,Country_New, by="country")
head(Df)

```
<br>
<br>
<br>

### 3) Build a linear regression model
```{r}

### Transform the data type
Df$GDP.growth.rate <- as.integer(Df$GDP.growth.rate)
Df$Unemployment <- as.integer(Df$Unemployment)
Df$Education.expenditure <- as.integer(Df$Education.expenditure)


### Build a Linear regression model with 9variables to predict the Happiness.Score
mod1 <- lm(Happiness.Score~Economy+Life.Expectancy + Generosity + Government.Corruption +GDP.growth.rate + Services.and.other.activity + Education.expenditure +using.the.Internet + Unemployment, data= Df)
summary(mod1)
```
#### We build a linear regression model predicting the Happiness.Score. We selected 9 important variables, and we have a R suqare of 0.8062. 
<br>
<br>
<br>

### 4) Happiness.score analysis in Aisa only
```{r}
Regions <- unique(Df$Region)
Regions

Df %>% 
  filter(grepl("Asia", Region, ignore.case = TRUE))%>%
  arrange(desc(Happiness.Score))
```
#### From the table we can see that Israel has the higest Happiness.Score, and Yemen has the lowest score

<br>
<br>
<br>


### 5) Happiness.score in differen Regions around the world
```{r}
Region_stats <- 
  Df%>%
  group_by(Region)%>%
  summarise(avg_score = mean(Happiness.Score))%>%
  arrange(desc(avg_score))
Region_stats


filter(Region_stats, avg_score == min(avg_score))

```
#### From the table we can conclude that people living in NorthernAmerican have higest happiness.score, and people living in MiddleAfrica have the lowest happiness.score
<br>
<br>
<br>

### 6) User-written function
```{r}
Check <- function(country_name){
  information = filter(Df, country == country_name)
  Region = information$Region
  Happiness.Rank = information$Happiness.Rank
  Happiness.Score = information$Happiness.Score

  table <- c(Happiness.Rank, Happiness.Score)
  names(table) = c("Happiness.Rank", "Happiness.Score")
  return(table)
  

}


Check('China')
```
#### 1) We build a functon to take the country names as inputs. Our function can check a countriy's Happiness.Rank and Happiness.Score.
#### 2) For example, China ranks 79th, and has a happiness.score of 5.273.
<br>
<br>
<br>



#### Plots
<br>
<br>
<br>

#### 7) Freedom - Happiness plot

```{r}
cor(Happy$Happiness.Score, Happy$Freedom)
favstats(~Happiness.Score, data = Happy)

ggplot(data = Df, aes(x = Freedom, y = Happiness.Score)) + 
  geom_point()+
  aes(color=Region) +
  xlab ('Freedom') + 
  ylab ('Happiness Score')+ 
  ggtitle("Freedom -- Happiness plot")

```
#### From the graph we can claim that freedom score and happniess score have a positive relationship

<br>
<br>

#### 8) Economy - Happiness plot

```{r}
ggplot(data = Df, aes(x = Economy, y = Happiness.Score)) + 
  geom_point()+
  aes(size=using.the.Internet) + 
  stat_smooth() + 
  xlab ('Economy_score') + 
  ylab ('Happiness Score')+ 
  ggtitle("Economy -- Happiness plot")
```
#### From the plot we can conclude that economy score and happniess score have a positive relationship.
<br>
<br>

#### 9) Life.Expectancy - Happiness plot
```{r}
ggplot(data = Df, aes(x = Life.Expectancy, y = Happiness.Score)) + 
  geom_point()+
  aes(size=Generosity) + 
  stat_smooth() + 
  xlab ('Life.Expectancy') + 
  ylab ('Happiness Score')+ 
  ggtitle("Life.Expectancy -- Happiness plot")
```
#### From the plot we can conclude that Life.Expectancy score and happniess score have a positive relationship.

